Towards Bringing Together Numerical Methods for Partial Differential Equation and Image Processing

### 2020-03-31 paper/paper.tex
Wraping things up

The simulations were performed numerically by solving the differential equation describing the flow --- the Navier-Stokes equation. This was done with a numerical solver library --- \emph{HiFlow}~\cite{gawlok2017} ---  that works on the base of the Finite element method~\cite{strang2008analysis}. The solver supports parallelization with \emph{MPI}~\cite{mpi} and \emph{Open}MP~\cite{dagum1998openmp} and we used 12 MPI processes to run each simulation. For all of the simulations, the timestep for of solver was set to $0.035$ seconds. This means a single time step of the simulations corresponds to a $0.035$ seconds of physical time.

@@

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_x}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_y}
  \end{subfigure}
  \begin{center}
    Constant model
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_speed_x_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_speed_y_1}
  \end{subfigure}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_speed_x_2}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_speed_y_2}
  \end{subfigure}
  \begin{center}
    Images form the inflow speed model. The upper two are from simulation with $g=1.3$ and the lower ones form simulation with $g=0.1$
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_fluid_x_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_fluid_y_1}
  \end{subfigure}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_fluid_x_2}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_fluid_y_2}
  \end{subfigure}
  \begin{center}
  Images form the viscosity-density model. The upper two are from simulation with $\nu=1.3, \rho=2.3$ and the lower ones form simulation with $\nu=1.3, \rho=2.3$
  \end{center}

@@

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i190}
  \end{subfigure}  
  \caption{The simulation that starts at the 10.\@ frame exhibits unintuitive results. This happens because the PSNR is based on MSE and it can go up if the network produces a homogeneous black image. In this case the PSNR can get higher in the second of the predicted frames. This, however, does not meant that the visual fidelity of the frames is good.}\label{fig:rec_const_psnr}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i1436}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the inflow speed model on simulation with $g=1.5$. To note is that this simulations was part of the training set. This can explain why the PSNR drops slightly faster on the next charts when it comes to the simulation starting at the 120.\@frame.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the inflow speed model on simulation with $g=3.9358$.
  \end{center}
  \caption{In the evaluations of the inflow speed models, we consistently saw the same behavior as in the constant models. The simulations that at are started at a frame with smaller index are of a lower quality.}\label{fig:rec_speed_psnr}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=4.1764$ and $\nu=0.001023$.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=5.7647$ and $\nu=0.001517$.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=6.8235$ and $\nu=0.002876$.
  \end{center}  
  \caption{All of the simulations for the charts are from the test set of the viscosity-density model. The evaluations with starting frame of 0 are consistent with the results from the other models. The evaluations starting at frame 120 are, however, much more stable as indicated by the standard deviation. We again attribute that to the over all stability of the model due to the large amount of train data used to train the models of this type.}\label{fig:rec_fluid_psnr}
\end{figure}

@@

@@

We believe that out approach can be extended to other simulation types as well. We see our work as preliminary research and we believe that the method can be further developed and investigated. Strategies for improving the models that were considered but not implemented include:
\begin{itemize}
\item[$\cdot$] using an LSTM~\cite{lstm} layer in the networks --- our models have the limitation that during training they don't know about the artifact on the images caused by themselves. We think an LSTM layer can mitigate this problem.
\item[$\cdot$] preprocessing the input for and postprocessing the output of the networks --- morphological operations like opening and closing can also be used to minimize the artifacts in the predicted frames.
\item[$\cdot$] further investigation of the effects of using RGB-images instead of grayscale ones.
\end{itemize}
\clearpage
\begin{figure}[H]
  \vspace*{-2.5cm}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/sims/constant/x_step_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/sims/constant/y_step_1}
  \end{subfigure}

@@

  \caption{These frames are generated by the constant model. The artifacts become apparent after the 10.\@ one.}\label{fig:const_sim}
\end{figure}
\vspace*{-1cm}
\begin{figure}[H]
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/sims/fluid/x_step_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/sims/fluid/y_step_1}
  \end{subfigure}

>>>>

The simulations were performed numerically by solving the differential equation describing the flow --- the Navier-Stokes equation. This was done with a numerical solver library --- \emph{HiFlow}~\cite{gawlok2017} ---  that works on the base of the Finite element method~\cite{strang2008analysis}. The solver supports parallelization with \emph{MPI}~\cite{mpi} and \emph{OpenMP}~\cite{dagum1998openmp} and we used 12 MPI processes to run each simulation. For all of the simulations, the timestep for of solver was set to $0.035$ seconds. This means a single time step of the simulations corresponds to a $0.035$ seconds of physical time.

@@

  \begin{subfigure}{.5\textwidth}
    \centering
    \large{x-velocity}
    \includegraphics[width=1\linewidth]{imgs/single_constant_x}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \large{y-velocity}
    \includegraphics[width=1\linewidth]{imgs/single_constant_y}
  \end{subfigure}
  \begin{center}
    Constant model
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/x_recursive_0_speed_8}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/y_recursive_0_speed_8}
  \end{subfigure}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/x_recursive_0_speed_39}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/y_recursive_0_speed_39}
  \end{subfigure}
  \begin{center}
    Images form the inflow speed model. The upper two are from simulation with $g=1.9487$ and the lower ones form simulation with $g=3.9358$
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/x_recursive_0_fluid_25}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/y_recursive_0_fluid_25}
  \end{subfigure}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/x_recursive_0_fluid_300}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/y_recursive_0_fluid_300}
  \end{subfigure}
  \begin{center}
  Images form the viscosity-density model. The upper two are from simulation with $\rho=4.1764, \nu=0.001023$ and the lower ones form simulation with $\rho=4.1764, \nu=0.001023$.
  \end{center}

@@

\begin{figure}[H]
  \vspace*{-4cm}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i190}
  \end{subfigure}  
  \caption{The simulation that starts at the 10.\@ frame exhibits unintuitive results. This happens because the PSNR is based on MSE and it can go up if the network produces a homogeneous black image. In this case the PSNR can get higher in the second of the predicted frames. This, however, does not meant that the visual fidelity of the frames is good.}\label{fig:rec_const_psnr}
\end{figure}
\vspace*{-1cm}
\begin{figure}[H]
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i1436}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the inflow speed model on simulation with $g=1.5$. To note is that this simulations was part of the training set. This can explain why the PSNR drops slightly faster on the next charts when it comes to the simulation starting at the 120. \@frame.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the inflow speed model on simulation with $g=3.9358$.
  \end{center}
  \caption{In the evaluations of the inflow speed models, we consistently saw the same behavior as in the constant models. The simulations that at are started at a frame with smaller index are of a lower quality.}\label{fig:rec_speed_psnr}
\end{figure}
\vspace*{-1cm}
\begin{figure}[H]
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=4.1764$ and $\nu=0.001023$.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=5.7647$ and $\nu=0.001517$.
  \end{center}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s120}
  \end{subfigure}
  \begin{center}
    Recursive evaluations of the viscosity-density model on simulation with $\rho=6.8235$ and $\nu=0.002876$.
  \end{center}  
  \caption{All of the simulations for the charts are from the test set of the viscosity-density model. The evaluations with starting frame of 0 are consistent with the results from the other models. The evaluations starting at frame 120 are, however, much more stable as indicated by the standard deviation. We again attribute that to the over all stability of the model due to the large amount of train data used to train the models of this type.}\label{fig:rec_fluid_psnr}
\end{figure}

@@

\noindent\textbf{Performance:} All of the models were evaluated on an Nvidia GTX 980Ti GPU.\@In all cases, the forward propagation of a single simulation frame took between 5 and 20 ms (milliseconds). In comparison, during the data generation, HiFlow needed between 1000 and 1500 ms for a single simulation timestep while using 12 MPI processes. This does not include the time needed for rendering the results of the simulation. To note is that this is not a one-to-one comparison as HiFlow runs on the CPU and uses MPI and OpenMP for parallelization. We, however, think that the results show the potential of our approach when it comes to performance.

@@

We believe that out approach can be extended to other simulation types as well. We see our work as preliminary research and we believe that the method can be further developed and investigated. Strategies for improving the models that were considered but not implemented include:
\begin{itemize}
\item[$\cdot$] using an LSTM~\cite{lstm} layer in the networks --- our models have the limitation that during training they don't know about the artifact on the images caused by themselves. We think an LSTM layer can mitigate this problem.
\item[$\cdot$] preprocessing the input for and postprocessing the output of the networks --- morphological operations like opening and closing can also be used to minimize the artifacts in the predicted frames.
\item[$\cdot$] further investigation of the effects of using RGB-images instead of grayscale ones.
\end{itemize}
\begin{figure}[H]
  \vspace*{-4.5cm}
  \begin{subfigure}{.5\textwidth}
    \centering
    \large{x-velocity}
    \includegraphics[width=1\linewidth]{imgs/sims/constant/x_step_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \large{y-velocity}
    \includegraphics[width=1\linewidth]{imgs/sims/constant/y_step_1}
  \end{subfigure}

@@

  \caption{These frames are generated by the constant model. The artifacts become apparent after the 10.\@ one.}\label{fig:const_sim}
\end{figure}
\vspace*{-1cm}
\begin{figure}[H]
  \begin{subfigure}{.5\textwidth}
    \centering
    \large{x-velocity}
    \includegraphics[width=1\linewidth]{imgs/sims/fluid/x_step_1}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \large{y-velocity}
    \includegraphics[width=1\linewidth]{imgs/sims/fluid/y_step_1}
  \end{subfigure}
